Dependent Object Types

### 2012-09-17 doc/fool/article.tex
rev

We propose DOT as a new type-theoretic foundation of Scala and
languages like it. The properties we are interested in modeling are
Scala's path-dependent types and abstract type members, as well as its
mixture of nominal and structural typing. Compared to previous
approaches~\cite{nuObj,FS}, we make no attempt to model inheritance or
mixin composition. Indeed we will argue that such concepts are better
modeled in a different setting.

@@

Compared to Scala, DOT also simplifies type members. In DOT, a type
member is declared by its lower and upper bounds. Furthermore, unlike
Scala, there is no notion of uninstantiated type member. Here is an
example inspired by~\cite{vt}:
\begin{lstlisting}
trait Food
trait Animal {
  type Meal <: Food
  def eat(meal: Meal) {}
}
\end{lstlisting}
In Scala, \code{Meal} is an uninstantiated type member of \code{Animal}:
it is not possible to instantiate an \code{Animal} without providing a
concrete binding for \code{Meal}. In DOT, \code{Meal} is a type member
with a lower bound of $\Bot$ and an upper bound of \code{Food}. It is
possible to instantiate an \code{Animal} without further specifying
\code{Meal}, though it would be not possible to feed it. Now,
we define \code{Cow} by refining \code{Animal}.
\begin{lstlisting}
trait Grass extends Food
trait Cow extends Animal {
  type Meal = Grass
}
\end{lstlisting}
In Scala, the type alias \code{type Meal = Grass} defines a concrete
binding for \code{Meal}. Now, we can instantiate \code{Cow}s and feed
them \code{Grass}. In DOT, such a type alias is declared by giving it
identical lower and upper bounds.

>>>>

We propose DOT as a new type-theoretic foundation of Scala and
languages like it. The properties we are interested in modeling are
Scala's path-dependent types and abstract type members, as well as its
mixture of nominal and structural typing through the use of refinement
types. Compared to previous approaches~\cite{nuObj,FS}, we make no
attempt to model inheritance or mixin composition. Indeed we will
argue that such concepts are better modeled in a different setting.

@@

Compared to Scala, DOT also simplifies type members. In DOT, a type
member is declared by its lower and upper bounds. Furthermore, unlike
Scala, there is no notion of uninstantiated type member. Here is an
example inspired by~\cite{vt}:
\begin{lstlisting}
trait Food
trait Animal {
  type Meal <: Food
  def eat(meal: Meal) {}
}
\end{lstlisting}
In Scala, \code{Meal} is an uninstantiated type member of \code{Animal}:
it is not possible to instantiate an \code{Animal} without providing a
concrete binding for \code{Meal}. In DOT, \code{Meal} is a type member
with a lower bound of $\Bot$ and an upper bound of \code{Food}. It is
possible to instantiate an \code{Animal} without further specifying
\code{Meal}, though it would not be possible to feed it. Now,
we define \code{Cow} by refining \code{Animal}.
\begin{lstlisting}
trait Grass extends Food
trait Cow extends Animal {
  type Meal = Grass
}
\end{lstlisting}
In Scala, the type alias \code{type Meal = Grass} defines a concrete
binding for \code{Meal}. Now, we can instantiate \code{Cow}s and feed
them \code{Grass}. In DOT, such a type alias is declared by giving it
identical lower and upper bounds.
